---
title: "horror"
author: "Jon Minton"
date: "2022-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Links/background 

- [Link to the page](https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-11-01)
- [Link to the data description](https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-11-01#data-dictionary)

## Packages and data 

```{r}
library(tidyverse)
dta <- read_csv("https://github.com/rfordatascience/tidytuesday/raw/master/data/2022/2022-11-01/horror_movies.csv")


```

## Inspect 

Any quirks/errors etc? 
```{r}
dta
```
Issues 

- missing labelled as 0 for `budget`, `revenue`, `runtime`
- adult: the adult flag not described in the github page 

Opportunities/questions

- runlength: Has there been any trend towards higher or lower run length over time? 
- linked genres: what are the most and least common genres that are paired with horror? - profitability: What's the relationship between budget and return? 
    - modelling of profitability: what variables explain higher returns (relative or absolute)? 


More advanced possibilities:

- inflation adjustment: use another package to pull inflation rates and inflation adjust both budgets and returns
- key terms: apply something like term document frequency approach to the description. Are implied topic memberships predictive of either genres or return?  
- image analysis: There are links to the posters. I remember seeing a data science example where the average hue/colour wheel for different posters were calculated for different film genres. (As expected horror and thrillers tended to use darkest average hues)



## Question one: has runlength changed over time? 

```{r}
dta %>% 
  filter(runtime != 0) %>% 
  ggplot(aes(release_date, runtime)) + 
  geom_point()

```

There are a small number of exceptionally long horror films. What are they?  

```{r}
dta %>% 
  filter(runtime > 600)
```

These appear to in effect be TV series released as films 

And there are some films that are long but not as long

```{r}
dta %>% 
  filter(between(runtime, 300, 600)) 
```

Again, these appear to be series. 

I'm going to look at films up to 300 minutes, and use some translucency to better check for a trend 

```{r}
dta %>% 
  filter(between(runtime, 1, 299)) %>% 
  ggplot(aes(release_date, runtime)) + 
  geom_point(alpha = 0.15)
```

It looks like the average length of horror films increased slightly over the 60s, but has remained in the 90-100 min range ever since. 

Since the 2000s it appears two things have happened:

- the variation in feature length films has increased slightly
- there has been an increase in short films. Perhaps this is due to the internet and easier distribution of independent content? 

The bifurcation in film lengths since around 2000 means fitting a single smoother line to the data appears a bad idea. 

Let's look at some examples of short films since around 2000

```{r}
dta %>% 
  filter(between(runtime, 1, 50)) %>% 
  filter(release_date >= lubridate::ymd("2000-01-01")) %>% 
  sample_n(30)
```

Many of these do not have a budget or revenue listed against them, suggesting they were not officially released. Perhaps many of these were 'calling cards' for new directors/talent etc? 

Some of these are documentaries. I wonder if some of these short horror films could in fact be short 'making of' features released alongside DVD releases of full length horror films? 

We can perhaps test this by looking for 'Documentary' in genre_names

```{r}
dta %>% 
  filter(between(runtime, 1, 400)) %>% 
  mutate(is_documentary = str_detect(genre_names, "Documentary")) %>% 
  ggplot() +
  ggridges::geom_density_ridges(aes(runtime, is_documentary), alpha = 0.5) 

```
As the above function uses a density estimator it's technically wrong as it has nonzero density for negative runtimes. It suggests that there's more variation in runtime for documentaries than other genres, however. 

## Return on investment 

What's the relationship between budget and revenue? 

```{r}
dta %>% 
  filter(budget > 0, revenue > 0) %>% 
  ggplot() + 
  geom_point(aes(x = budget, y = revenue)) 
```

No obvious relationship. Does a double log scale help? 


```{r}
dta %>% 
  filter(budget > 0, revenue > 0) %>% 
  ggplot() + 
  geom_point(aes(x = budget, y = revenue)) +
  scale_y_log10() + 
  scale_x_log10() 

```
Now the relationship is clearer. Let's tidy this up a bit more 

```{r}

bks <- c(
  1e2, 1e3, 
  1e4, 1e5,
  2e5, 5e5,
  1e6, 2e6,
  5e6, 1e7,
  2e7, 5e7,
  1e8, 2e8, 5e8
)

lbls <- c(
  "100", "1k",
  "10k", "100k",
  "200k", "500k",
  "1M", "2M",
  "5M", "10M",
  "20M", "50M",
  "100M", "200M", "500M"
)

dta %>% 
  filter(budget > 0, revenue > 0) %>% 
  ggplot(aes(x = budget, y = revenue)) + 
  geom_point(alpha = 0.2) +
  scale_y_log10(breaks = bks, labels = lbls) + 
  scale_x_log10(breaks = bks, labels = lbls) +
  coord_fixed() + 
  stat_smooth(method = "lm") + 
  geom_abline(slope = 1, intercept = 0, colour = "red", linetype = "dashed") + 
  labs(x = "Budget ($)", y = "Revenue ($)",
       title = "Budget against revenues, log-log scale",
       subtitle = "Regardless of budget size, horror films on average make money",
       caption = "Source: https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-11-01") +
  ggdark::dark_theme_grey() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```
The predicted line is above the equality line (red dashed line). So, the average revenue of horror films is greater than the average budget. 

I guess this is why horror films are made! 


### Modelling 

We know there's an above-zero return on investment, but what factors push this return up or down? 

```{r}
dta_prep <- dta %>% 
  filter(budget > 0, revenue > 0, runtime > 0) 

mod_00 <- lm(log(revenue) ~ log(budget), dta_prep)

df_preds <- tibble(
  budget = bks,
  pred_revenue = exp(predict(mod_00, newdata = tibble(budget = bks)))
) %>% 
  mutate(
    roi = (pred_revenue - budget) / budget
  )

df_preds
```

This suggests that low budget (\$1M) films can be expected to get around a 70% ROI, falling to around 60% at \$10M, to around 45% at \$100M. 

What's the median budget for a horror film (considering only last 20 years)?


```{r}
dta_prep %>% 
  filter(release_date >= lubridate::dmy("01-01-2002")) %>% 
  summarise(
    med_budget = median(budget),
    lower_q_budget = quantile(budget, 0.25),
    upper_q_budget = quantile(budget, 0.75)
  )

```

So the median horror film costs around \$6M, and half of all films have budgets between \$1M and around \$17M. 


Anyway, what other factors affect the predicted revenue? 

```{r}

 
mod_01 <-   lm(log(revenue) ~ log(budget) + runtime, dta_prep)
mod_02 <-   lm(log(revenue) ~ log(budget) + poly(runtime, 2), dta_prep)
mod_03 <-   lm(log(revenue) ~ log(budget) + poly(runtime, 3), dta_prep)

anova(mod_00, mod_01, mod_02, mod_03)
```

This suggests runtime linearly increases predicted revenue (though should be careful with extrapolating to very long runtimes). 

```{r}
mod_04 <- lm(log(revenue) ~ log(budget) + runtime + is_not_english, 
             dta_prep %>% mutate(is_not_english = original_language != "en"))

anova(mod_01, mod_04)
```

Non-English films have lower predicted returns, but this isn't quite statistically significant. 

```{r}
mod_05 <- lm(log(revenue) ~ log(budget) + runtime + release_date, dta_prep)
anova(mod_01, mod_05)
```

Adding a time term does seem to improve fit. However polynomials can't be calculated directly with datetime objects. Maybe they can if just the year and month component is included

```{r}
mod_06 <- lm(
  log(revenue) ~ log(budget) + runtime + poly(year, 2), 
  dta_prep %>% 
    mutate(year = lubridate::year(release_date) + lubridate::month(release_date) / 12)
  )

anova(mod_05, mod_06)
```
So a polynomial on time does seem to improve fit. 

Let's now consider a couple more potential variables: 

- Whether it's part of a collection
- Whether it's a pure horror or mixed genre

```{r}
mod_07 <- lm(
  log(revenue) ~ log(budget) + runtime + poly(year, 2) + is_collection, 
  dta_prep %>% 
    mutate(
      year = lubridate::year(release_date) + lubridate::month(release_date) / 12,
      is_collection = !is.na(collection_name)
      )
  )
summary(mod_07)
# So, it appears being part of a collection/franchise increases revenues
anova(mod_06, mod_07)

```

```{r}

mod_08 <- lm(
  log(revenue) ~ log(budget) + runtime + poly(year, 2) + is_collection, 
  dta_prep %>% 
    mutate(
      year = lubridate::year(release_date) + lubridate::month(release_date) / 12,
      is_collection = !is.na(collection_name)
      )
  )

dta_prep %>% 
  mutate(
    year = lubridate::year(release_date) + lubridate::month(release_date) / 12,
    is_collection = !is.na(collection_name)
  ) %>% 
  mutate(
    num_additional_genres = stringr::str_count(genre_names, ","),
    horror_first = stringr::str_extract(genre_names, "^[A-Za-z]*") == "Horror",
    mixed_genres = num_additional_genres > 0
  ) %>% 
  lm( 
    log(revenue) ~ log(budget) + runtime + poly(year, 2) + is_collection + mixed_genres, 
    data = . 
  ) -> mod_09
summary(mod_09)
anova(mod_08, mod_09)
# Looks like mixed_genres doesn't add to model fit

dta_prep %>% 
  mutate(
    year = lubridate::year(release_date) + lubridate::month(release_date) / 12,
    is_collection = !is.na(collection_name)
  ) %>% 
  mutate(
    num_additional_genres = stringr::str_count(genre_names, ","),
    horror_first = stringr::str_extract(genre_names, "^[A-Za-z]*") == "Horror",
    mixed_genres = num_additional_genres > 0
  ) %>% 
  lm( 
    log(revenue) ~ log(budget) + runtime + poly(year, 2) + is_collection + horror_first, 
    data = . 
  ) -> mod_10
summary(mod_10)
anova(mod_08, mod_10)
# But knowing something is first and foremost a horror genre does predict 
# revenue


dta_prep %>% 
  mutate(
    year = lubridate::year(release_date) + lubridate::month(release_date) / 12,
    is_collection = !is.na(collection_name)
  ) %>% 
  mutate(
    num_additional_genres = stringr::str_count(genre_names, ","),
    horror_first = stringr::str_extract(genre_names, "^[A-Za-z]*") == "Horror",
    mixed_genres = num_additional_genres > 0
  ) %>% 
  lm( 
    log(revenue) ~ log(budget) + runtime + poly(year, 2) + is_collection + horror_first + num_additional_genres, 
    data = . 
  ) -> mod_11
summary(mod_11)
anova(mod_10, mod_11)

# Number of additional genres doesn't have a statistically signifant effect on 
#log revenues

# So, our best model is mod_10

#    log(revenue) ~ log(budget) + runtime + poly(year, 2) + is_collection + horror_first
```

### Predictions 

Having iterated through a few model specifications, we can use our best model to project revenues for a number of different scenarios. Let's consider the following:

- Budget: \$1M, \$5M, \$10M, \$20M
- runtime: 80m, 90m, 100m
- year: 1985, 1995, 2005, 2015
- collection: TRUE, FALSE
- horror_first: TRUE, FALSE


```{r}
df_scenarios <- expand_grid(
  budget = c(1e6, 5e6, 1e7, 2e7),
  runtime = c(80, 90, 100),
  year = c(1985, 1995, 2005, 2015),
  is_collection = c(TRUE, FALSE),
  horror_first = c(TRUE, FALSE)
)

predictions <- 
  bind_cols(
    df_scenarios,
    predicted_revenue = predict(mod_10, newdata = df_scenarios) %>% exp()
  ) %>% 
    mutate(
      predicted_profit = predicted_revenue - budget,
      predicted_roi = predicted_revenue / budget
    )

predictions
```

There are a lot of scenarios here. So perhaps we should create a little app for allowing the user to select between them? 

```{r}
write_rds(predictions, "data/scenarios_and_predictions.rds")
write_rds(predictions, "show_predictions/scenarios_and_predictions.rds")


```

